\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}

\title{Mathematical Logic Exam}
\author{Liu Siyu}
\date{December 2023}

\begin{document}

\maketitle

\begin{enumerate}
    \item Function $f$ : $A \rightarrow B$ is ( \textbf{surjective/onto} ) if for every element of B, 
        it has some preimage in A.
    \item A set is said to be ( \textbf{countable} ) iff there is a one-to-one mapping 
        from it to the set of natural numbers. 
    \item The set of infinite binary sequences is ( \textbf{uncountable} ).
    \item if A is countable and non-empty then the set of all finite
        sequences of members of A is ( \textbf{countable} ).    
    \item Formalize the following statement into a well-formed formula (wff): \\
        ``If the computer system doesn't work well, then either the software has some bugs, or the hardware is not functioning.'' \\
        Let $A_1$ denote ``the computer system works well'', let $A_2$ denote ``the software has some bugs'', 
        let $A_3$ denote ``the hardware is functioning''. \\
        \bm{$((\lnot A_1)\rightarrow (A_2 \lor (\lnot A_3)))$}
    \item The possible lengths of a wff is $n$, $n$ is a positive integer, then $n$ can't be 
        ( \textbf{2, 3, 6} )
    \item A truth assignment $v$ satisfies the wff $(\dots (A_1 \leftrightarrow A_2)\leftrightarrow \dots \leftrightarrow A_n)$
        iff the number of $i$ that $v(A_i) = F$ is $n$, $1 \leq i \leq n$, and ( \textbf{ n is even} ).
    \item Of the following three sets of wffs, list the label(s) of satisfiable ones: \\
        (1). $\{\lnot A_3, A_1 \leftrightarrow A_3\} \quad$ (2). $\emptyset  \quad$ 
        (3). $\{A_1, A_1\rightarrow A_2, A_2\rightarrow (A_1 \land \lnot A_1)\}$ \\
        (4). $\{((\lnot A_1 \land A_2)\rightarrow (A_1 \rightarrow A_2))\}$ \\
        \textbf{1, 2, 4}
    \item Let $\alpha$ be the wff $((A_2\rightarrow (A_1\rightarrow A_3)) \leftrightarrow ((A_2 \land A_1)\rightarrow A_3))$,
        the truth assignment for the sentence symbols is $v(A_1)=T,\; v(A_2)=T,\; v(A_3)=F$, then 
        $\overline{v}(\alpha)$ is ( \textbf{T} ).
    \item Write a formula that is tautologically equivalent to $(\lnot((\lnot (A_1 \lor A_2)) \rightarrow A_3))$
        whose only connectives are $\lnot$ and $\land$. \\
        \bm{$(((\lnot A_1)\land (\lnot A_2))\land (\lnot A_3))$}
\end{enumerate}

\begin{enumerate}
\item Which of the following statements is false? (C)
    \begin{enumerate}
        \item If A and B are countable, $A\times B$ is countable.
        \item An empty set is countable.
        \item Let S denote the set of infinite binary sequences, then S is countable.
        \item if A is countable and non-empty then the set of all finite sequences of members of A is countable.
    \end{enumerate}

\item Which of the following statements about wff is false? (C)
    \begin{enumerate}
    \item Any expression with more left parentheses than right parentheses is not a wff.
    \item Suppose that $\alpha$ is a wff not containing the negation symbol ¬, the length of $\alpha$ is odd.
    \item Suppose that $\alpha$ is a wff not containing the negation symbol ¬, it's possible that less than a quarter of the symbols are sentence symbols.
    \item $(A_1 \land A_2 \land A_3)$ is not a wff.
    \end{enumerate}

\item Which of the following statements doesn't hold in sentential logic? (D)
    \begin{enumerate}
        \item $\{A, (\lnot A)\}\models B$.
        \item $\{A, (A\rightarrow B) \}\models B$.
        \item $\emptyset \models (((A \land B)\rightarrow C)\rightarrow D)\rightarrow ((A\rightarrow C)\rightarrow D)$.
         \item $(A\leftrightarrow (B\leftrightarrow C))\models ((A\land (B\land C))\lor ((\lnot A)\land ((\lnot B)\land (\lnot C)))$.
    \end{enumerate}

\item Which of the sets of sentential connectives are complete? (A)\\
(1). $\{\land, \lor, \lnot \}$ \\
(2). $\{\land, \rightarrow \}$ \\
(3). $\{\land, \lor, \rightarrow, \leftrightarrow \}$ \\
(4). $\{\lnot, \land \}$
    \begin{enumerate}
        \item (1)(4)
        \item (1)(3)
        \item (1)(3)(4)
        \item (1)(2)(3)(4)
    \end{enumerate}

\item Which of the following statements are true? (D)\\
(1). If $\alpha \in \Sigma$, then $\Sigma \models \alpha$.\\
(2). If $\Sigma \models \alpha$, then there is a finite $\Sigma_0 \subseteq \Sigma$ such that $\Sigma_0 \models \alpha$.\\
(3). If $\Sigma \models \alpha$ and $\alpha \models \beta$, then $\Sigma \models \beta$.\\
(4). $\Sigma \not\models \alpha$ iff $\Sigma \cup \{\lnot \alpha \}$ is satisfiable.\\
    \begin{enumerate}
        \item (1)(2)(3)
        \item (1)(2)(4)
        \item (2)(3)(4)
        \item (1)(2)(3)(4)
    \end{enumerate}

\item Which of the following wffs are sentences? (B)\\
(1). $(\forall xAx\rightarrow Bx)$\\
(2). $(\forall x(Px\rightarrow \forall xQx ))$\\
(3). $(\forall x(Py\rightarrow \exists y Q x y))$\\
(4). $(\forall x \forall y (P x y \rightarrow \exists z Qxz))$
    \begin{enumerate}
        \item (1)(4)
        \item (2)(4)
        \item (1)(2)(4)
        \item (3)(4)
    \end{enumerate}

\item Which of the following wff is not valid? (B)
    \begin{enumerate}
        \item $\forall x (\dot{P}x\rightarrow \exists y\dot{P}y)$
        \item $\dot{P}xy \rightarrow \exists y \dot{P}yy$
        \item $\forall x(\dot{P}x\rightarrow \dot{Q}x)\land \dot{Q}y \rightarrow \dot{P}y$
        \item $\forall x(\dot{P}x\rightarrow \dot{Q}x)\land \exists y \dot{P}y \rightarrow \exists y \dot{Q}y$
    \end{enumerate}

\end{enumerate}

\end{document}
